---
title: "part1"
output: pdf_document
---

### Question 1

Download file `gene_experession.tsv` from github, read file csv, make row name is column one and print first 6 genes.

```{r}
download.file("https://github.com/markziemann/SLE712_files/raw/master/bioinfo_asst3_part1_files/gene_expression.tsv", "gene_expression.tsv")
df <- read.csv('gene_expression.tsv', sep='\t', stringsAsFactors = FALSE, row.names = 1)
head(df, 6)
```

Try to access a gene by gene name.

```{r}
df['ENSG00000223972', ]
```

Make a new column is mean of other columns

```{r}
df$mean <- rowMeans(df[, 1:2])
head(df)
```

### Question 3

Create sorted dataframe by mean column order
```{r}
sorted_df <- df[order(df$mean), ]
```

Show top 10-highest mean genes
```{r}
top10genes <- row.names(tail(sorted_df, 10))
top10genes
```

### Question 4
Take mean column after compare this column with 10 (<10). The resutl will be a list of boolean array. Number of genes with mean lower than 10 is sum of this list.

```{r}
number_genes <- sum(df$mean < 10)
number_genes
```

### Question 5
```{r}
hist(df$mean, xlab='mean of gene')
```


### Question 6
Download growth data from github and load it into a dataframe

```{r}
download.file("https://github.com/markziemann/SLE712_files/raw/master/bioinfo_asst3_part1_files/growth_data.csv", "growth_data.csv")
df <- read.csv('growth_data.csv')
head(df)
```
Print column names of dataframe

```{r}
colnames(df)
```
Mean and standard deviation at 2004 (start)

```{r echo=T}
mean_2004 <- mean(df$Circumf_2004_cm)
cat('Mean at 2004: ', mean_2004)
cat('\n')
sd_2004 <- sd(df$Circumf_2004_cm)
cat('SD at 2004: ', sd_2004)
```

Mean and standard deviation at 2019 (end)
```{r}
mean_2019 <- mean(df$Circumf_2019_cm)
cat('Mean at 2019: ', mean_2019)
cat('\n')
sd_2019 <- sd(df$Circumf_2019_cm)
cat('SD at 2019: ', sd_2019)
```

